---
title: "lab2_analysis"
author: "Nicole Linares"
date: "2026-09-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Logistic regression modeling 10-year coronary heart disease (CHD) risk from established risk factors (age, blood pressure, cholesterol, smoking, diabetes), using the Framingham Heart Study teaching dataset.

```{r load-data}
url <- "https://raw.githubusercontent.com/GauravPadawe/Framingham-Heart-Study/master/framingham.csv"
fram <- read.csv(url)

dim(fram)
head(fram)
```

## Handling Missing Data

Same approach as the Python notebook: several variables have missing values (glucose: 388, education: 105, BPMeds: 53, totChol: 50, cigsPerDay: 29, BMI:19, heartRate: 1). Since only 13.7% of rows have any missing value and the data are evenly distributed, I dropped incomplete rows via complete-case analysis (dropping missing value rows) rather than imputation (estimation). Worth noting as a limitation.

```{r missing-data}
colSums(is.na(fram))

fram_clean <- na.omit(fram)
dim(fram_clean)
```

## Logistic Regression Model

Predicting 10-year CHD risk from five established risk factors: age, systolic blood pressure, total cholesterol, current smoking status, and diabetes. Logistic regression is a logical choice because the outcome (TenYearCHD) is binary (0/1). Linear regression is not appropriate for a binary outcome.

```{r fit-model}
model <- glm(TenYearCHD ~ age + sysBP + totChol + currentSmoker + diabetes, data = fram_clean, family = binomial)
summary(model)
```

```{r odds-ratios}
odds_ratios <- exp(coef(model))
conf_int <- exp(confint(model))

results_table <- cbind(`Odds Ratio` = odds_ratios, conf_int)
round(results_table, 3)

```

