---
title: "lab2_analysis"
author: "Nicole Linares"
date: "2026-09-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Logistic regression modeling 10-year coronary heart disease (CHD) risk from established risk factors (age, blood pressure, cholesterol, smoking, diabetes), using the Framingham Heart Study teaching dataset.

```{r load-data}
url <- "https://raw.githubusercontent.com/GauravPadawe/Framingham-Heart-Study/master/framingham.csv"
fram <- read.csv(url)

dim(fram)
head(fram)
```

## Handling Missing Data

Same approach as the Python notebook: several variables have missing values (glucose: 388, education: 105, BPMeds: 53, totChol: 50, cigsPerDay: 29, BMI:19, heartRate: 1). Since only 13.7% of rows have any missing value and the data are evenly distributed, I dropped incomplete rows via complete-case analysis (dropping missing value rows) rather than imputation (estimation). Worth noting as a limitation.

```{r missing-data}
colSums(is.na(fram))

fram_clean <- na.omit(fram)
dim(fram_clean)
```

## Logistic Regression Model

Predicting 10-year CHD risk from five established risk factors: age, systolic blood pressure, total cholesterol, current smoking status, and diabetes. Logistic regression is a logical choice because the outcome (TenYearCHD) is binary (0/1). Linear regression is not appropriate for a binary outcome.

```{r fit-model}
model <- glm(TenYearCHD ~ age + sysBP + totChol + currentSmoker + diabetes, data = fram_clean, family = binomial)
summary(model)
```

```{r odds-ratios}
odds_ratios <- exp(coef(model))
conf_int <- exp(confint(model))

results_table <- cbind(`Odds Ratio` = odds_ratios, conf_int)
round(results_table, 3)

```

## Visualization: Predicted CHD Risk by Age and Smoking Status

Utilizes the fitted model to plot predicted 10-year CHD risk across the age range, comparing current smokers vs. non-smokers. Blood pressure and cholesterol are held at the sample mean, and diabetes at 0 (meaning non-diabetic), so the plot isolates the effect of age and smoking status specifically.

```{r prediction-data}
age_range <- seq(min(fram_clean$age), max(fram_clean$age))
mean_sysBP <-mean(fram_clean$sysBP)
mean_totChol <- mean(fram_clean$totChol)

pred_smoker <- data.frame(age = age_range, sysBP = mean_sysBP, totChol = mean_totChol, currentSmoker = 1, diabetes = 0)
pred_non_smoker <- pred_smoker 
pred_non_smoker$currentSmoker <- 0 

risk_smoker <-predict(model, newdata = pred_smoker, type = "response")
risk_nonsmoker <- predict(model, newdata = pred_non_smoker, type = "response")
```

```{r plot-risk}
plot(age_range, risk_smoker, type = "l", col = "firebrick", lwd = 2,
     xlab = "Age (years)", ylab = "Predicted 10-year CHD risk",
     main = "Predicted 10-Year CHD Risk by Age and Smoking Status",
     ylim = c(0, max(risk_smoker, risk_nonsmoker)))
lines(age_range, risk_nonsmoker, col = "steelblue", lwd = 2)
legend("topleft", legend = c("Current smoker", "Non-smoker"),
       col = c("firebrick", "steelblue"), lwd = 2, title = "Smoking Status")
grid()
```

## Interpretation

Based on the results of the logistic regression and the odds ratio, I chose to represent smoking status in a graph to show the risk of smoking and CHD across the age span. The odds ratio was 1.67x for smoking and 1.07x for age, making them strong predictors. Despite smoking status, age seemed to be the stronger predictor of developing CHD. The smokers had a significantly higher risk as well, but smoking alone was not the biggest contributor to CHD. Compared to the plot in the Jupyter notebook, the scale is smaller so it seems that the smoker and non smoker correlations are even less, but adjusting for scale to make both the same would likely end up plotting the same result. The outcomes of the logistic regression and odds ratio were essentially identical.
